---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# goldfish.latent

<!-- badges: start -->
<!-- badges: end -->

`goldfish.latent` extends the `goldfish` package with models that include
latent variables.
Therefore, how to define the event sequence data objects for modeling and
the effects available for modeling uses `goldfish`'s data objects definitions
and statistics effects.
Inferences of the proposed models are made using Hamiltonian Chain Monte Carlo
as implemented in [Stan](https://mc-stan.org/).

## Installation

You can install the development version of goldfish.latent like so:

``` r
remotes::install_github("snlab-ch/goldfish.latent", build_vignettes = TRUE)
```

## Example

The first model introduced in `goldfish.latent` is the DyNAM with random effects.
Model formulation allows to include monadic statistic or actors' covariates
to explain the variability of the random effects.

```{r example, message=FALSE, results='hide'}
library(goldfish.latent)

# using cmdstanr for getting posterior samples and 
library(cmdstanr)

# using goldfish social evolution data set
library(goldfish)
data("Social_Evolution")
callNetwork <- defineNetwork(nodes = actors, directed = TRUE) |>
 linkEvents(changeEvent = calls, nodes = actors)
callsDependent <- defineDependentEvents(
 events = calls, nodes = actors, defaultNetwork = callNetwork
)
data2stan <- CreateData(
 randomEffects = list(inertia ~ 1),
 fixedEffects = callsDependent ~ recip + trans
)

stanCode <- CreateModelCode(data2stan)

mod01 <- cmdstan_model(stanCode)
mod01Samples <- mod01$sample(
  data = data2stan[["dataStan"]],
  parallel_chains = 4, chains = 4,  iter_warmup = 500, iter_sampling = 500,
  show_messages = FALSE
)

```

Using `cmdstanr` functionalities makes easier fit to use summary and plotting
function from packages that works with MCMC posterior samples.
For the summary of the posterior samples `cmdstanr` uses the 
[`posterior`](https://mc-stan.org/posterior/) package.
Plots from the posterior samples can be done using, for example,
[`bayesplot`](https://mc-stan.org/bayesplot/) package. 

```{r}
mod01Samples$summary("beta")

# names fixed effects coincides with colnames(data2stan$dataStan$X)

# Plots from the posterior distribution can be generated using `bayesplot`
# or any other package that handle MCMC posteriors
```


## Code of Conduct

Please note that the `goldfish.latent` project is released with a
[Contributor Code of Conduct](https://contributor-covenant.org/version/2/1/CODE_OF_CONDUCT.html).
By contributing to this project, you agree to abide by its terms.
