---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# goldfish.latent

<!-- badges: start -->
<!-- badges: end -->

`goldfish.latent` extends the `goldfish` package with models that include latent
variables.
Therefore, how to define the event sequence data objects for modeling and
the effects available for modeling uses `goldfish` data objects definitions and
statistics effects.
Inferences of the proposed models are done using Hamiltonian Chain Monte Carlo as
implemented in [Stan](https://mc-stan.org/). 

## Installation

You can install the development version of goldfish.latent like so:

``` r
remotes::install_github("snlab-ch/goldfish.latent", build_vignettes = TRUE)
```

## Example

The first model introduced in `goldfish.latent` is the DyNAM with random effects.
Model formulation allows to include monadic statistic or actors' covariates to explain
the variability of the random effects.

```{r example}
library(goldfish.latent)

# using goldfish social evolution data set
library(goldfish)
data("Social_Evolution")
callNetwork <- defineNetwork(nodes = actors, directed = TRUE) |>
 linkEvents(changeEvent = calls, nodes = actors)
callsDependent <- defineDependentEvents(
 events = calls, nodes = actors, defaultNetwork = callNetwork
)
data2stan <- CreateData(
 randomEffects = list(inertia ~ 1),
 fixedEffects = callsDependent ~ recip + trans
)
```

## Code of Conduct

Please note that the `goldfish.latent` project is released with a
[Contributor Code of Conduct](https://contributor-covenant.org/version/2/1/CODE_OF_CONDUCT.html).
By contributing to this project, you agree to abide by its terms.
